### 2011-05-22 paper/discussion.tex
added figures to tex.  win!

\begin{itemize}
\item what about the nonparametric version of this?
\end{itemize}