### 2011-05-22 paper/discussion.tex

\begin{itemize}
\item what about the nonparametric version of this?
\item discuss the broader goal, from the beginning of search engine time, to combine categorization and free text.  this, to our knowledge, is the first principled approach to doing so
\end{itemize}