### 2011-05-26 paper/discussion.tex
Updated Experiment section and other minor things

We have described a mixed membership model with hierarchical supervision. We have demonstrated this model in the context of document modeling with hierarchical multi-label supervision. Such a model is appropriate in domains where there are hierarchical constraints among the labels such as is the case in an IS-A hierarchy.

...

\begin{itemize}
\item what about the nonparametric version of this?
\item discuss the broader goal, from the beginning of search engine time, to combine categorization and free text.  this, to our knowledge, is the first principled approach to doing so
\end{itemize}