### 2011-05-29 paper/main.tex

\documentclass{article}
\usepackage{nips10submit_e,times}
%\documentstyle[nips07submit_09,times]{article}
\usepackage[square,numbers]{natbib}
\usepackage{amsmath, epsfig}
\usepackage{amsfonts}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{easybmat}
\usepackage{footmisc}
\usepackage{lscape}
\renewcommand\algorithmiccomment[1]{// \textit{#1}}
%
\newcommand{\ignore}[1]{}
\newcommand{\comment}[1]{}
\DeclareMathOperator*{\argmax}{arg\,max}

\title{Hierarchically Supervised Latent Dirichlet Allocation}
%Supervised Topic Modeling in Clinical Text}

\author{
Adler Perotte\hspace{1cm} Nicholas Bartlett \hspace{1cm} Noemie Elhadad \hspace{1cm} Frank Wood\\
Columbia University, New York, NY 10027, USA \\
\texttt{\{ajp9009@dbmi,bartlett@stat,noemie@dbmi,fwood@stat\}.columbia.edu}
%\texttt{pfau@neurotheory.columbia.edu} 
%\texttt{\{bartlett,fwood\}@stat.columbia.edu} 
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
\newcommand{\X}{\mathcal{X}}

\nipsfinalcopy

\begin{document}

\maketitle

\begin{abstract}
\input{abstract}
\end{abstract}

\section{Introduction}
\label{sec:introduction}
\input{introduction}

%\section{Background}

\subsection{Related Work}

\label{sec:related_work} Latent dirichlet allocation (LDA) is a generative
probabilistic model of corpora that represents documents as a mixed
membership bag-of-words. Also known as topic models, these models
infer the latent structure, or topics, of documents in a corpus. Each
document is represented as a collection of words, generated from a
set of topic assignments (one for each word), where each topic assignment
is drawn from a distribution over topics \citep{Blei2003}.

Supervised latent Dirichlet allocation (sLDA) builds on LDA by incorporating
an exponential family response variable. Although there are many models
for making predictions based on free text, sLDA is unique in that
it is a generative model, it represents documents as a mixed-membership,
and constrains the inference of the latent structure of the documents
by its predictability of the response variable. In other words, sLDA
infers topics such that the model is capable of a high predictive
likelihood for words in a document and the response variable associated
with a document. This approach has been shown to outperform both LASSO
(L1 regularized least squares regression) and LDA followed by least
squares regression \citep{BleiMcAuliffe2008}.

There have been many models that incorporate both latent models of
text and some form of supervision \citep{Ramage2009,DiscLDA,wangbleifeifei08,RelationalLDA}.
One set of models that are particularly relevant to HSLDA are Chang
and Blei's hierarchical models for document networks (Relational Topic
Models). In that family of models, they encountered a similar scenario
where the lack of a link did not truly indicate absence. In hierarchically
labeled data, negative labels are uncommon and the lack of a label
in the hierarchy is not equivalent to a negative label. Therefore,
as in the work of Chang and Blei, we employ regularization to account
for the lack of negative labels. This will be discussed further in
\ref{sec:model}.

% move to related work?
While there have been some attempts to automatically classify groups
of patients as a potential preliminary step to ICD-9 code assignment
\citep{Ruch2008,FreitasJunior2006,RibeiroNeto2001,Brown2006}, fully
automatic assignment of ICD-9 codes to medical text became a more
prevalent research topic only in the last few years. A subset of earlier
work proposed various methods on small corpora, based on a few specific
diseases \citep{Rao2003} but the most recent and promising work on
the subject was inspired by the 2007 Medical NLP Challenge: {}``International
Challenge: Classifying Clinical Free Text Using Natural Language Processing\textquotedblright{}
(website). Most of the classification strategies included word matching
and rule-based algorithms. \citep{Goldstein2007,Crammer2007,Farkas2008}.
The data set given to the participants consisted only of documents
that were 1-2 lines each and all of the documents were radiology reports
- clearly limiting the scope of potential ICD-9 codes which could
be assigned. The only paper which has attempted to work with a document
scope as large as ours was the 2008 Lita et al publication \citep{Lita2008}.
Lita proposed support vector machine and bayesian ridge regression
methods to assign appropriate labels to the documents but did not
utilize the ICD-9 hierarchy to leverage more comprehensive predictions.

\section{Model}

\label{sec:model} We define here a hierarchically supervised LDA
model. Although we will focus on document modeling in our description
and experiments, this model applies equally well to other collections
of discrete data with hierarchically constrained labels.

We assume a pre-specified set of labels $\mathcal{L}=\left\{ l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $.
Each document is assigned a response of either -1 or 1 for at least
one, but potentially many labels in $\mathcal{L}$. The label, $l$,
for a document, $d$, will be used interchangeably to refer to the
observed response of document $d$ to label $l$. The label set is
assumed to be structured as an {}``is-a'' hieararchy.
To understand this, consider a hierarchy where label $l_{1}$ is a
parent of label $l_{2}$. If document $d$ has a positive response
to label $l_{2}$ then it will also have a positive response to label
$l_{1}$. Conversely, if document $d$ has a negative response to
label $l_{1}$ then it will also has a negative response to $l_{2}$.
To capture this hierarchical structure we model the labeling of documents
using a generative cascade of conditional probit regression models.

% generalize to arbitrary codes (not just ICD-9's
%
\begin{figure}[h]
%tbp] %  figure placement: here, top, bottom, or page
 \centering \includegraphics[scale=0.3]{Graphical_Model2} \caption{adapted sLDA model}

\label{fig:example} 
\end{figure}

The fixed parameters of the model are the number of topics $K$, the
number of unique words in the vocabulary $V$, the number of documents
$D$, as well as the mean, $\mu$, and the standard deviation, $\sigma$,
used in a normal prior distribution. The hyper-parameters $\alpha^{\prime}$,
$\alpha$, and $\gamma$ are weight parameters for Dirichlet prior
distributions. We will denote the $K$-dimensional Dirichlet distribution
as Dir$_{K}(\cdot)$ and the $K$ dimensional normal distribution
as $\mathcal{N}_{K}(\cdot)$.

We will now describe the stochastic generative process which defines
our model. The graphical model is show in Figure~\ref{fig:example}.
%Given the number of topics, K, and broad gamma priors on hyperparameters,
%the generative process is as follows:

\begin{enumerate}
\item For each topic $k=1,\ldots,K$:

\begin{itemize}
\item Draw a distribution over words $\mathbf{\phi_{k}}\sim{\rm Dir}_{V}\left(\gamma\mathbf{1}\right)$,
where $\mathbf{1}$ is a vector of ones of length $V$ 
\end{itemize}
\item For each label $l\in\mathcal{L}$:

\begin{itemize}
\item Draw the regression coefficients $\mathbf{\eta_{l}}\mid\sigma\sim\mathcal{N}_{K}\left(\mu I_{K},\sigma I_{K}\right)$,
where $I_{K}$ is the $K$ dimensional identity matrix 
\end{itemize}
\item Draw a prior over topic proportions $\mathbf{\beta}\mid\alpha'\sim{\rm Dir}_{K}\left(\alpha^{\prime}\mathbf{1}\right)$
\item For each document $d=1,\ldots,D$:

\begin{itemize}
\item Draw topic proportions $\mathbf{\theta_{d}}\mid\beta,\alpha\sim{\rm Dir}_{K}\left(\alpha\mathbf{\beta}\right)$ 
\item For $n=1,\ldots,N_{d}$:

\begin{itemize}
\item Draw topic assignment $z_{n,d}\mid\mathbf{\theta_{d}}\sim{\rm Multinomial}\left(\mathbf{\theta_{d}}\right)$ 
\item Draw word $w_{n,d}\mid z_{n,d},\mathbf{\beta_{1:K}}\sim{\rm Multinomial}\left(\mathbf{\beta_{z_{n,d}}}\right)$ 
\end{itemize}
\item For each label $l\in\mathcal{L}$:

\begin{itemize}
\item Draw $a_{l,d}\mid z_{1:N_{d},d},\eta_{l},y_{parent(l),d}\sim\begin{cases}
\mathcal{N}\left(\bar{z}^{T}\eta_{l},1\right), & y_{parent(l)}=1\\
\mathcal{N}\left(\bar{z}^{T}\eta_{l},1\right)I\left(a_{l,d}<0\right), & y_{parent(l)}=-1\end{cases}$ where $\bar{z_{d}}=N_{d}^{-1}\sum_{n=1}^{N}z_{n,d}$ %\item Draw $a_{l, d} \ | \ z_{1:N_d,d}, \eta_l \sim \mathcal{N} \left(\bar z_d^{T} \eta_{l},1\right)$, where $\bar z_d=N_d^{-1}\sum_{n=1}^{N_d}z_{n,d}$

\item Set the response variable \[
y_{l,d}\mid a_{l,d}=\begin{cases}
\ \ \ 1 & \text{if \ensuremath{a_{l,d}>0}and \ensuremath{y_{{\rm parent}(l),d}=1}}\\
-1 & \text{otherwise}\end{cases}\]

\end{itemize}
\end{itemize}
\end{enumerate}
This type of generative model is known as a probit regression model.
Probit regression models are a type of discriminative probabilistic
model similar to logistic regression. However, instead of using the
logistic sigmoid as the link function, the probit regression model
uses the CDF for a standard normal distribution - the inverse of which
is known as the probit. In this case, the regression is conditional
on the parents according to the constraints of the labeling hierarchy.
The latent variables $a_{l,d}$ utilized here are also known as an
auxiliary variables because the are introduced to make exact Gibbs
sampling possible and are not of primary interest.

Given that negative labels are uncommon and that the absence of a
label is not equivalent to a negative label, we apply an informative
prior to the regression parameters, $\mathbf{\eta}_{\mathcal{L}}$,
in the form of a negative prior that encodes a bias towards being
truly negative in the absence of a label.

\section{Inference}

\label{sec:inference} In the Bayesian approach to statistical modeling,
the primary task of inference is to find the posterior distribution
over the unobserved parameters of the model. However, it is often
possible and desirable to integrate over certain variables in the
model, also known as collapsing. In our model, it will often be the
case that the set of labels $\mathcal{L}$ is not fully observed for
every document. We will define $\mathcal{L}_{d}$ to be the subset
of labels which have been observed for document $d$. It is straightforward
to integrate out the variables $a_{l^{\prime},d}$ and $y_{l^{\prime},d}$
for $l^{\prime}\in\mathcal{L}\backslash\mathcal{L}_{d}$ from the
full generative model. We can also integrate out the parameters $\mathbf{\phi_{1:K}}$
and $\mathbf{\theta_{1:D}}$ as in \citet{Griffiths04}. Therefore,
in our model the latent variables are $\mathbf{z}=\{z_{1:N_{d},d}\}_{d=1,\ldots,D},\mathbf{\eta}=\{\mathbf{\eta_{l}}\}_{l\in\mathcal{L}},\mathbf{a}=\{a_{l^{\prime},d}\}_{l^{\prime}\in\mathcal{L}_{d},d=1,\ldots,D},\mathbf{\beta},\alpha,\alpha^{\prime}$
and $\gamma$.

The posterior distribution we seek cannot be solved in closed form.
This is often the case in evaluating posterior distributions of non-trivial
probabilistic models. We will appeal to one of the common methods
for approximating posterior distributions in the face of intractable
normalization factors: Markov chain Monte Carlo (MCMC) sampling. Since
in this model it is possible to sample from the conditional distributions
for all variables we will use the Gibbs sampling algorithm to obtain
an approximation to this posterior. %In particular, we will derive a collapsed Gibbs sampler.
%Given an observation of a set of observed labels and a document, the posterior distribution for the latent variables is
%\begin{equation}
%p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma\mid w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)\label{eq:Posterior}\end{equation}
%\[
%=\frac{p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma,w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)}{\int_{\theta,\phi,a,\eta,\alpha,\alpha',\beta,\gamma}\sum_{z}p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma,w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)}\]

\subsection{Gibbs Sampler}

We derive a collapsed Gibbs sampler for this model by considering
the individual conditional probability distributions for each of the
unobserved variables. We use the notation $\mathbf{z}_{-(n,d)}$ to
denote $\mathbf{z}\backslash z_{n,d}$. %, marginalizing
%$\theta_{1:D}$ and $\phi_{1:K}$. For details regarding collapsing
%in LDA models see \citet{Griffiths04}.

\subsubsection{$p\left(z_{n,d}\mid\mathbf{z_{-\left(n,d\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)$}

First we consider the conditional distribution of the assignment variable
for each word $n=1,\ldots,N_{d}$ in documents $d=1,\ldots,D$. The
conditional distribution does not include $\theta_{1:D}$ and $\phi_{1:K}$
because they have been integrated out as in the collapsed Gibbs sampler
\citep{Griffiths04}. The conditional distribution of $z_{n,d}$ is
proportional to the joint distribution of its markov blanket. %\begin{equation}
%p\left(z_{d,n}\mid\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\propto p\left(z_{d,n},\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\end{equation}
%Due to the factorization of this model, we can rewrite the joint distribution
%as the following:
\begin{equation}
p\left(z_{n,d}\mid\mathbf{z_{-\left(n,d\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\propto\prod_{l\in\mathcal{L}_{d}}p\left(a_{l,d}\mid\mathbf{z},\eta_{l}\right)p\left(z_{n,d}\ |\ \mathbf{z_{-\left(n,d\right)}},\mathbf{a},\mathbf{w},\alpha,\beta,\gamma\right).\end{equation}
 The product is only over the subset of labels $\mathcal{L}_{d}$
which have been observed for document $d$. By isolating terms that
depend on $z_{n,d}$ and absorbing all other terms into a normalizing
constant as in \citep{Griffiths04} we find \begin{eqnarray}
\lefteqn{p\left(z_{n,d}=k\mid\mathbf{z_{-\left(n,d\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\propto}\nonumber \\
 & \hspace{2cm}\left(n_{\left(\cdot\right),d}^{k,-\left(n,d\right)}+\alpha\beta_{k}\right)\frac{n_{w_{n,d},\left(\cdot\right)}^{k,-\left(n,d\right)}+\gamma}{\left(n_{\left(\cdot\right),\left(\cdot\right)}^{k,-\left(n,d\right)}+V\gamma\right)}\prod_{l\in\mathcal{L}_{d}}\exp\left\{ -\frac{\left(\bar{z}_{d}^{T}\eta_{l}-a_{l,d}\right)^{2}}{2}\right\} .\label{eq:z-likelihood}\end{eqnarray}
 Here, $n_{v,d}^{k,-\left(n,d^{\prime}\right)}$ represents the number
of words of type $v$ in document $d$ assigned to topic $k$ omitting
the $n^{{\rm th}}$ word of document $d^{\prime}$. The notation $(\cdot)$
in the subscript means the count resulting from summing over the omitted
subscript variable. Given Equation \ref{eq:z-likelihood}, $p\left(z_{d,n}\mid\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\mathbf{\beta},\gamma\right)$
can be sampled through enumeration.

\subsubsection{$p\left(\mathbf{\eta_{l}}\mid\mathbf{z},\mathbf{a},\sigma\right)$}

We now consider the conditional distribution of the regression coefficients
$\mathbf{\eta_{l}}$ for $l\in\mathcal{L}$. Given that $\mathbf{\eta_{l}}$
and $a_{l,d}$ are distributed normally, the posterior distribution
of $\mathbf{\eta_{l}}$ is normally distributed with mean $\hat{\mu}$
and covariance $\hat{\mathbf{\Sigma}}$ such that % (probably not the right place for this) We evaluated the model over various values of $\sigma$ where $\sigma=\left\{ 0.01,0.1,0.25,1,2\right\} $.
\begin{eqnarray}
\hat{\mathbf{\Sigma}}^{-1} & = & \mathbf{I}\sigma^{-1}+\bar{\mathbf{Z}}^{T}\bar{\mathbf{Z}}\\
\hat{\mu} & = & \hat{\mathbf{\Sigma}}_{i}\left(\mathbf{1}\frac{\mu}{\sigma}+\bar{\mathbf{Z}}^{T}\mathbf{a}_{l}\right).\end{eqnarray}
 This is a standard result from normal Bayesian linear regression
\citep{BishopPRML}. Here, $\bar{\mathbf{Z}}$ is a $D\times K$ matrix
such that row $d$ of $\mathbf{\bar{Z}}$ is $\bar{z}_{d}$, and $\mathbf{a}_{l}=[a_{l,1},a_{l,2},\ldots,a_{l,D}]^{T}$.
%p\left(\eta_{i_{l,c}}\mid\mathbf{z}_{1:D},\mathbf{a};\sigma\right)=\mathcal{N}\left(\eta_{i_{l,c}}\mid\hat{\mu}_{i},\hat{\Sigma}_{i}\right)\end{equation}
%\[
%\[

\subsubsection{$p\left(a_{l,d}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)$}

%and \textmd{$p\left(y_{m,i}\mid\mathbf{a}\right)$}}
The auxiliary variables $a_{l,d}$ must be sampled for documents $d=1,\ldots,D$
and $l\in\mathcal{L}_{d}$. The conditional posterior distribution
of $a_{l,d}$ is the truncated normal distribution \begin{equation}
p\left(a_{l,d,}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)\propto\frac{1}{\sqrt{2\pi}}exp\left\{ -\frac{1}{2}\left(a_{l,d}-\mathbf{\eta_{l}}^{T}\mathbf{\bar{z}}_{d}\right)\right\} I\left(a_{l,d}y_{l,d}>0\right).\end{equation}
 This conditional distribution can be sampled using an inverse CDF
method. %However, if $y_{d,i_{l,c}}$ is unobserved then $a_{d,i_{l,c}}$must
%be sampled jointly with $y_{d,i_{l,c}}$ to ensure that the Markov
%chain is ergodic. Suppose that $a_{d,i_{l,c}}$ is sampled to have
%a negative value and $y_{d,i_{l,c}}$ is apporopriately sampled at
%-1. Although there exist valid states where $a_{d,i_{l,c}}>0$ and
%$y_{d,i_{l,c}}=1$, they will never be reached by such a Markov chain
%since $p\left(a_{d,i_{l,c}}<0\mid y_{d,i_{l,c}}=-1\right)=1$ and
%$p\left(y_{d,i_{l,c}}=-1\mid a_{d,i_{l,c}}<0\right)=1$. Therefore,
%to ensure ergodicity, $a_{d,i_{l,c}}$and $y_{d,i_{l,c}}$ must be
%sampled from the joint distribution as shown in Equation \ref{eq:Probit-Joint}.
%\begin{equation}
%p\left(a_{d,i_{l,c}},y_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y}_{-\left(d,i_{l,c}\right)},\mathbf{\eta}\right)\propto p\left(y_{i_{l,c}}\mid\mathbf{a},\mathbf{y}_{-\left(l,c\right)}\right)p\left(a_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)\end{equation}
%\begin{equation}
%p\left(y_{i_{l,c}}\mid\mathbf{a},\mathbf{y}_{-\left(l,c\right)}\right)=\delta\left(sign\left(a_{d,i_{l,c}}\right)=y_{i_{l,c}}\right)p\left(y_{i_{l,c}}\mid y_{parents_{l,c}}\right)\prod_{i_{\hat{l},\hat{c}}\in children_{l,c}}p\left(y_{i_{\hat{l},\hat{c}}}\mid y_{i_{l,c}}\right)\end{equation}
%\begin{equation}
%p\left(y_{i_{l,c}}=-1\mid y_{parent{}_{l,c}}\right)=\begin{cases}
%1, & y_{parent_{l,c}}=-1\\
%0.5, & y_{parent_{l,c}}=1\end{cases}\end{equation}
%\[
%p\left(a_{d,i_{l,c}},y_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y}_{-\left(d,i_{l,c}\right)},\mathbf{\eta}\right)\]
%\begin{equation}
%=\begin{cases}
%\mathcal{N}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)p\left(y_{d,i_{l,c}}\mid a_{d,i_{l,c}}\right), & y_{parent_{l,c}}=1,\forall y_{i_{\hat{l},\hat{c}}}\in y_{children_{l,c}},y_{i_{\hat{l},\hat{c}}}=-1\\
%trunc\mathcal{N}^{-}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)\delta\left(y_{d,i_{l,c}}=-1\right), & y_{parent_{l,c}}=-1\\
%trunc\mathcal{N}^{+}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)\delta\left(y_{d,i_{l,c}}=1\right), & \exists y_{i_{\hat{l},\hat{c}}}\in y_{children_{l,c}}\setminus y_{i_{\hat{l},\hat{c}}}=1\\
%0 & otherwise\end{cases}\end{equation}
%where $\mathbf{Y}_{-\left(d,i_{l,c}\right)}$ denotes all of the response
%variables excluding the response variable being sampled.

\subsubsection{$p\left(\beta\mid\mathbf{z},\alpha^{\prime},\alpha\right)$}

In our model, we place a hierarchical Dirichlet prior over topic assignments.
This flexible distribution allows for an asymmetric prior over document
level distributions over topics \citep{WallachMiMc2009}. This prior
shares many features with the hierarchical Dirichlet process and inference
over this distribution proceeds in a very similar fashion.

Posterior inference is performed using the {}``direct assignment''
method of \citet{TehJorBea2006}. \begin{equation}
\beta\sim{\rm Dir}\left(m_{\left(\cdot\right),1}+\alpha^{\prime},m_{\left(\cdot\right),2}+\alpha^{\prime},\ldots,m_{\left(\cdot\right),K}+\alpha^{\prime}\right)\end{equation}
 \begin{equation}
p\left(m_{d,k}=m\mid\mathbf{z},\mathbf{m}_{-\left(d,k\right)},\beta\right)=\frac{\Gamma\left(\alpha\beta_{k}\right)}{\Gamma\left(\alpha\beta_{k}+n_{d,k}\right)}s\left(n_{d,k},m\right)\left(\alpha\beta_{k}\right)^{m}\end{equation}
 where $s\left(n,m\right)$ represents stirling numbers of the first
kind.

\subsubsection{$p\left(\alpha\right)$, $p\left(\alpha'\right)$, $p\left(\gamma\right)$}

The hyperparameters $\alpha$, $\alpha^{\prime}$, and $\gamma$ are
given broad ${\rm Gamma}(1,1000)$ prior distributions and sampled
via the Metropolis-Hastings algorithm.

\section{Experiments}

\label{sec:experiments}

%\begin{itemize}
%\item Describe Data for both Experiments
%\item Describe the Four different conditions
%\item Describe how they are all evaluated, with and without hierarchy
%\end{itemize}
In the section we describe the application of HSLDA for prediction
in two hierarchically structured domains. Firstly, we describe using
discharge summaries to predict diagnoses, encoded as ICD-9 codes.
Discharge summaries are documents that are authored by clinicians
to summarize the course of a hospitalization. ICD-9 codes are used
mainly for billing purposes to indicate the conditions for which a
patient was treated. Secondly, we describe using Amazon.com product
descriptions to predict product categories.

\subsection{Data and Pre-Processing}

\subsubsection{Diagnosis Prediction}

Our data set was gathered from the clinical data warehouse of NewYork-Presbyterian
Hospital. The data consisted of free-text discharge summaries and
their respective ICD-9 codes. A discharge summary is a clinical report
prepared by a physician or other health professional at the conclusion
of a hospital stay or series of treatments. The note outlines the
patient's chief complaint, diagnostic findings, therapy administered,
patient's response to the chosen therapy, the treatment plan and the
recommendations upon discharge. The ICD-9 codes used to structure
the discharge summary data are part of a controlled terminology which
is the international standard diagnostic classification for epidemiological,
health management, and clinical purposes. The ICD-9 codes are organized
in a rooted-tree structure, with each edge representing an is-a relationship
between parent and child, such that the parent diagnosis subsumes
the child diagnosis. For example, the code representing {}``Pneumonia
due to adenovirus'' is a child of the code representing {}``Viral
pneumonia'' where the former is a type of the latter. In the hospital,
ICD-9 codes are generated manually by trained medical coders, who
review all the information in the discharge summary.

The text of the discharge summaries were pre-processed such that each
document would be represented as counts over a 10,000 word vocabulary.
The Natural Language Toolkit was used to tokenize the text. A vocabulary
was identified by first sorting terms based on a global term frequency-inverse
document frequency measure. The top 10,000 words which were not identifying
in some way (a name, place, or identifying number) were selected for
inclusion in the vocabulary.

For each hospitalization there are usually several ICD-9 codes assigned
for billing purposes. These codes are known to be quite specific but
not very sensitive \citep{Birmetal2005}. Regardless of that fact,
this is one of the only sources for information on patient diagnoses
aside from the free text. %Aside from prediction, one of the goals is to compare the sensitivity of predictions from the HSLDA model in comparison to the codes in a case where a test closer to ground truth is available. For this we will compare whether predictions for the ICD-9 code associated with anemia are better predicted by HSLDA or by the ICD-9 codes. Anemia was chosen because hemoglobin values are readily available and the definition of anemia according the World Health Organization is approximately 12.5, with a threshold of 12 for women and 13 for men [citation].

We worked within the guidelines of the Health Insurance Portability
and Accountability Act (HIPAA), which protects patient privacy and
the security of potentially identifying medical material, known as
personal health information (PHI). HIPAA covers any information within
a medical record that was created, used, or disclosed during the course
of providing a health care service and that can be used to identify
an individual. This study was approved by the Institutional Review
Board.

\subsubsection{Product Category Prediction}

Data for these experiments were obtained partially from the Stanford
Network Analysis Platform (SNAP) Amazon product metadata dataset \citep{SNAP}
and partially directly from the the Amazon.com website \citep{AMAZON}.
The product ID's and categorizations were obtained from the SNAP dataset
and the product descriptions were obtained directly from the website.

We were able to deduce the structure of the hierarchy for the Amazon.com
products directly since all ancestors in the hierarchy were included
with each category label. For example, {}``DVD / Genres / Science
Fiction \& Fantasy / Classic Sci-Fi'' is a single product category
for the DVD, {}``The Time Machine''. Each product was labeled with
multiple categories.

The vocabulary for this experiment was created by including the most
frequent 30K words omitting stopwords.

\subsection{Comparison Models}

We applied HSLDA, along with three other closely related models, to
the clinical data and the retail product data. Specifically, we evaluate
models including sLDA with independent regressors (hierarchical constraints
on labels ignored), HSLDA fit by performing LDA followed by tree-conditional
regressions, and HSLDA fit with fixed random regression parameters.
We will refer to the three comparison models as the sLDA model, the
separate HSLDA model, and the random HSLDA model, respectively. These
models were chosen as to highlight performance in absence of hierarchical
constraints, the effect of the combined inference procedure, and regression
performance attributable solely to the hierarchical constraints.

We evaluated model performance for all three models with a range of
values for $\mu$ ($\mu\in\left\{ -3,-2.8,-2.6,\ldots,1\right\} )$,
the mean prior parameter for regression parameters.

\subsection{Evaluation}
%\begin{itemize}
%\item Talk about hierarchy and no hierarchy
%\item Make explicit calculation of sensitivity and 1-specificity
%\end{itemize}
For each dataset, a held out set of 1,000 documents and accompanying
labels were used for evaluation. The two main methods of evaluation
for the model are prediction and topic quality. To evaluate predictive
performance for all comparison models equivalently, each model was
evaluated with two methods. The first method augmented the observed
labels in the held out set as well as their ancestors in the hierarchy
as

\subsubsection{Prediction}

The two measures for predictive performance used here include the
true positive rate and the false positive rate. We evaluate model
performance on held out data. A more ideal evaluation of performance
would include a manually labeled hierarchy since it is well known
that ICD-9 codes have a relatively low sensitivity.

Performance was evaluated against $p\left(y_{l,\hat{d}}\mid w_{1:N,d}\right)$.

\subsubsection{Topic Quality and Character}

TBD

\section{Results}
\label{sec:results}

\begin{figure}[htbp]
\begin{center}
\subfigure[][\label{1a}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_ancestors}}
\subfigure[][\label{1b}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_leafs}}
\subfigure[][\label{1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
\subfigure[][\label{1d}]{\includegraphics[width=.49\textwidth]{figs/one_m_spec_comparison_leafs}}
\caption{Out-of-sample ICD-9 code prediction from patient free-text discharge records.  In all figures solid is HSLDA, dashed are independent regressors + sLDA (hierarchical constraints on labels ignored), dotted is HSLDA fit by running LDA first then running tree-conditional regressions, and dot-dashed is HSLDA fit with fixed random regression parameters.  Top row:  \subref{1a} includes ancestor prediction performance,  \subref{1b} results are for given (leaf) labels alone.  Bottom row: \subref{1c} are the sensitivity curves from \subref{1b} aligned on threshold value,  \subref{1d} are the 1-specificity curves from \subref{1b} aligned on threshold value.}
\label{fig:icd9}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\subfigure[][\label{2a}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_ancestors_amazon}}
\subfigure[][\label{2b}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_leafs_amazon}}
\subfigure[][\label{2c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs_amazon}}
\subfigure[][\label{2d}]{\includegraphics[width=.49\textwidth]{figs/one_m_spec_comparison_leafs}}
\caption{Out-of-sample Amazon product code predictions from product free-text descriptions.  In all figures solid is HSLDA, dashed are independent regressors + sLDA (hierarchical constraints on labels ignored), dotted is HSLDA fit by running LDA first then running tree-conditional regressions, and dot-dashed is HSLDA fit with fixed random regression parameters.  Top row:  \subref{2a} includes ancestor prediction performance,  \subref{2b} results are for given (leaf) labels alone.  Bottom row: \subref{2c} are the sensitivity curves from \subref{2b} aligned on threshold value,  \subref{2d} are the 1-specificity curves from \subref{2b} aligned on threshold value.}
\label{fig:icd9}
\end{center}
\end{figure}

%We ran the Gibbs sampler for two days on 500 documents with 20 topics to learn the sLDA model which we then followed by a prediction of ICD-9 codes for 100 documents. The prediction resulted in 38 documents with at least 1 ICD-9 code assignment while some had over 200 assignments.

%\begin{figure}[h]%tbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[width=4in]{number_of_codes_histogram} 
%   \caption{distribution of ICD-9 code assignments over all of the 100 documents predicted}
%   \label{fig:example}
%\end{figure}

%The accuracy of the prediction was tested against the actual ICD-9 codes in the EHR record for each discharge summary.  Different statistical measures were applied to asses the predictions.

%\begin{figure}[h]%tbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[width=4in]{signal_detection} 
%   \caption{sensitivity, specificity, precision, recall and F-measure.}
%   \label{fig:example}
%\end{figure}

%Although the sensitivty is not as high as we would have hoped, it is important to note that the sLDA was run on a very limited sample of documents (500) due to time constraints.  Additionally, the “gold standard” data used are actual hospital results which are very noisy given the incompleteness and introduction of human error.

\section{Discussion}
\label{sec:discussion}
\input{discussion}
%We have proposed a new method for prediction of ICD-9 codes given a medical discharge summary note.  Our method explicitly takes advantage of the ICD-9 hierarchical IS-A relationships as well as the word distributions in each document.  We used Gibbs sampling for a supervised latent dirichlet allocation model to learn a distribution of words over twenty topics and their appropriate ICD-9 codes (response variable). The model was then tested on 100 discharge summaries for which the ICD-9 diagnoses were known.  The amount of ICD-9 codes produced for some discharge summaries was quite high because of the way the model worked with the IS-A hierarchy, highlighting the completeness of ICD-9 coding when using this sLDA model.  Overall, the model had extremely good specificity while the sensitivity was lower than anticipated.  Given our limited set of documents, and very noisy gold standard it would not be accurate to discount the use of sLDA in medical document classification and automatic coding.
%In future work, we plan to significantly expand the set of documents used to train the model (up to 15,000), as well as run several different iterations to learn the optimal topic number.

%A more automated approach for coding patient visits with the ICD-9 hierarchy has great promise to  reduce health care costs and increase completeness of medical coded data.  We hope that this implementation of supervised LDA will serve as preliminary work for a much larger study with more documents in the future.

\begin{small} \bibliographystyle{plainnat} \bibliographystyle{plainnat}
\bibliographystyle{plainnat}
\bibliography{refs}

\end{small}

\end{document}