### 2011-05-31 paper/abstract.tex
updated the abstract -- is it better?

We introduce hierarchically supervised latent Dirchlet allocation (HSLDA), a model for hierarchically and multiply-labeled bag-of-word data.  Examples of such data include web pages and their placement in link directories, product descriptions and placement(s) in product hierarchies, and free-text clinical records and diagnosis codes assigned to them.
Out-of-sample label prediction is the primary goal of this work, but
improved lower-dimensional representations of the bag-of-words data is also of interest.  
%We define a model that uses probit regressors on a conditionally dependent label hierarchy tied to latent Dirichlet allocation (LDA).  
We find that using the signal from hierarchical labels substantially improves out-of-sample label prediction in comparison to other models that don't utilize the structure of the labels.   We demonstrate HSLDA on large-scale data from medical document labeling and retail product categorization tasks.  We show improved label prediction performance and evidence that the learned topics also improve.

%Topic models are unsupervised models well suited to data exploration and information retrieval.  Supervised topic models use side-information, labels and the like, to improve the learned representations.

%The benefits of supervision in topic modeling 
%Current medical record keeping technology relies heavily upon human
%capacity to effectively summarize and infer information from free-text
%physician notes. We propose a novel method to suggest diagnostic code
%assignment for patient visits, based upon narrative medical notes.
%We applied a supervised latent Dirichlet allocation model to a corpora
%of free-text medical notes from NewYork - Presbyterian Hospital to
%infer a set of specific ICD-9 codes for each patient note. Evaluation
%of the predictions were conducted by comparison to a gold-standard
%set of ICD-9s assigned to a set of patient notes.