### 2011-06-02 paper/discussion.tex
discussion tiny tiny tiny edits

%Despite the proliferation of LDA related models in recent years, one could
%argue that the true killer app for LDA has yet to be discovered.  LDA makes
%sense for information retrieval (IR) because  the learned latent topic vectors
%serve as good low-dimensional document representations, and, a posteriori
%similarity between such vectors can be used to compute the similarity of
%documents.  Unfortunately, computational considerations render this approach
%to IR impractical in most settings.

The SLDA model family, of which HSLDA is a member, can be understood in two
different ways. One way is to see it as a family of topic models that improve
on the topic modeling performance of LDA via the inclusion of observed
supervision. An alternative, complementary way is to see it as a set of models
that can predict labels for bag-of-word data. A large diversity of problems
can be expressed as label prediction problems for bag-of-word data. A
surprisingly large amount of that kind of data possess structured labels,
either hierarchically constrained or otherwise. That HSLDA directly addresses
this kind of data is a large part of the motivation for this work. That it
outperforms more straightforward approaches should be of interest to
practitioners.

Variational Bayes has been the predominant estimation approach applied to sLDA
models. Hierarchical probit regression makes for tractable Markov chain Monte
Carlo SLDA inference, a benefit that should extend to other sLDA models should
probit regression be used for response variable prediction there too.

The results in Figures~\ref{fig:1a} and \ref{fig:1b} suggest that in most cases
it is better to do full joint estimation of HSLDA.  An alternative
interpretation of the same results is that, if one is more sensitive to the
performance gains that result from exploiting the structure of the labels, then
one can, in an engineering sense, get nearly as much gain in label prediction
performance by first fitting LDA and then fitting a hierarchical probit
regression.  There are applied settings in which this could be advantageous.

Extensions to this work include unbounded topic cardinality variants and
relaxations to different kinds of label structure.  Unbounded topic cardinality
variants pose interesting inference challenges.  Utilizing different kinds of
label structure is possible within this framework, but requires relaxing some
of the simplifications we made in this paper for expositional purposes.

%We have described a mixed membership model with hierarchical supervision. We
%have demonstrated this model in the context of document modeling with
%hierarchical multi-label supervision. Such a model is appropriate in domains
%where there are hierarchical constraints among the labels such as is the case
%in an IS-A hierarchy.

%\begin{itemize}
%\item what about the nonparametric version of this?
%\item discuss the broader goal, from the beginning of search engine time, to combine categorization and free text.  this, to our knowledge, is the first principled approach to doing so
%\end{itemize}