### 2011-06-02 paper/experiments.tex
done

We applied HSLDA to data from two domains: predicting medical
diagnosis codes from hospital discharge summaries and predicting product
categories from Amazon.com product descriptions.

%In the section we describe the application of HSLDA for prediction
%in two hierarchically structured domains. Firstly, we describe using
%discharge summaries to predict diagnoses, encoded as ICD-9 codes.
%Discharge summaries are documents that are authored by clinicians
%to summarize the course of a hospitalization. ICD-9 codes are used
%mainly for billing purposes to indicate the conditions for which a
%patient was treated. Secondly, we describe using Amazon.com product
%descriptions to predict product categories.

\subsection{Data and Pre-Processing}

\subsubsection{Discharge Summaries and ICD-9 Codes}

Discharge summaries are authored by clinicians to summarize patient
hospitalization course. The summaries typically contain a record of patient
complaints, findings and diagnoses, along with treatment and hospital course.
For each hospitalization, trained medical coders review the information in the
discharge summary and assign a series of diagnoses codes. Coding follows the
ICD-9-CM controlled terminology, an international diagnostic classification for
epidemiological, health management, and clinical
purposes.\footnote{http://www.cdc.gov/nchs/icd/icd9cm.htm}  The ICD-9 codes are organized in a rooted-tree structure, with each
edge representing an is-a relationship between parent and child, such that the
parent diagnosis subsumes the child diagnosis. For example, the code for
{}``Pneumonia due to adenovirus'' is a child of the code for {}``Viral
pneumonia,'' where the former is a type of the latter.  It is worth noting that
the coding can be noisy. Human coders sometimes disagree~\cite{Challenge07},
tend to be more specific than sensitive in their
assignments~\cite{Birmetal2005}, and sometimes make
mistakes~\cite{FarzandipourEtAl10}.

The task of automatic ICD-9 coding has been investigated in the clinical
domain.  Methods range from manual
rules to online learning~\cite{Crammer2007,Goldstein2007,Farkas2008}.
Other work had leveraged larger datasets and experimented with K-nearest
neighbor, Naive Bayes, support vector machines, Bayesian Ridge Regression, as
well as simple keyword mappings, all with promising
results~\cite{LarkeyCroft95,RibeiroNeto2001,PakhomovEtAl06,Lita2008}.%Ruch2008 omitted

%Much of the work was triggered by the 2007 medical NLP community
%challenge~\cite{Challenge07}.
%The data in the challenge, however, differs from 
%ours in its scope. The datasets were smaller (1,000 training and 1,000 testing
%documents) and focused on radiology reports with a restricted number of ICD-9
%codes (45 of them, compared to 7K+ in our dataset).

Our dataset was gathered from the clinical data warehouse of
a large metropolitan hospital. It consists of 6,000 discharge summaries and
their associated ICD-9 codes (7,298 distinct codes overall), representing all
the discharges from the hospital in 2009. Summaries have 8.39 associated ICD-9
codes on average (std dev=5.01) and contain an average of 536.57 terms after
preprocessing (std dev=300.29). We split our dataset into 5,000 discharge
summaries for training and 1,000 for testing.

The text of the discharge summaries was tokenized with
NLTK.\footnote{http://www.nltk.org} A fixed vocabulary was formed by taking
the top 10,000 tokens with highest document frequency (exclusive of names,
places and other identifying numbers). The study was approved
by the Institutional Review Board and follows HIPAA (Health
Insurance Portability and Accountability Act) privacy guidelines.

%For each hospitalization there are usually several ICD-9 codes assigned
%for billing purposes. These codes are known to be quite specific but
%not very sensitive \citep{}. Regardless of that fact,
%this is one of the only sources for information on patient diagnoses
%aside from the free text. %Aside from prediction, one of the goals is to compare the sensitivity of predictions from the HSLDA model in comparison to the codes in a case where a test closer to ground truth is available. For this we will compare whether predictions for the ICD-9 code associated with anemia are better predicted by HSLDA or by the ICD-9 codes. Anemia was chosen because hemoglobin values are readily available and the definition of anemia according the World Health Organization is approximately 12.5, with a threshold of 12 for women and 13 for men [citation].

\subsubsection{Product Descriptions and Categorizations}

Amazon.com, an online retail store, organizes its catalog of products in a
mulitply-rooted hierarchy and provides textual product descriptions 
for most products. Products can be discovered by users
through free-text search and product category exploration. Top-level
product categories are displayed on the front page of the website and lower
level categories can be discovered by choosing one of the top-level categories.
Products can exist in multiple locations in the hierarchy.

In this experiment, we obtained Amazon.com product categorization data from 
the Stanford Network Analysis Platform (SNAP) dataset~\citep{SNAP}. 
Product descriptions were obtained separately from the
Amazon.com website directly. We limited our dataset to the collection of DVDs
in the product catalog.

%We were able to deduce the structure of the
%hierarchy for the Amazon.com products because all ancestors in the
%hierarchy were included with each category label. For example, {}``DVD / Genres
%/ Science Fiction \& Fantasy / Classic Sci-Fi'' is a single product category 
%for the DVD, {}``The Time Machine.''

Our dataset contains 15,130 product descriptions for training and 1,000
for testing. The product descriptions are shorter than the discharge summaries
(91.89 terms on average, std dev=53.08). Overall, there are 2,691 unique codes.
Products are assigned on average 9.01 codes (std dev=4.91). The vocabulary
consists of the most frequent 30,000 words omitting stopwords.

\subsection{Comparison Models}

We evaluated HSLDA along with two closely related models against
the two datasets. The comparison models included sLDA with independent
regressors (hierarchical constraints on labels ignored), HSLDA fit by first
performing LDA then fitting tree-conditional regressions. These models were
chosen to  highlight several aspects of HSLDA including performance in the
absence  of hierarchical constraints, the effect of the combined inference, and
regression performance attributable solely to the hierarchical constraints.

sLDA with independent regressors is the most salient comparison model
for our work. The distinguishing factor between HSLDA and sLDA is the
addition structure imposed on the label space, a distinction that we
hypothesized would result in a difference in predictive performance.

There are two components to HSLDA, LDA and a hierarchically constrained
response. The second comparison model is HSLDA fit by performing LDA first
followed by performing inference over the hierarchically constrained label
space. In this comparison model, the separate inference processes do not allow
the responses to influence the low dimensional structure inferred by LDA.
Combined inference has been shown to improve performance in sLDA
\citep{BleiMcAuliffe2008}. This comparison model examines not the structuring
of the label space, but the benefit of combined inference over both the
documents and the label space.

%The last comparison model is HSLDA with fixed and randomly selected regression
%parameters. There is a baseline benefit that the structure in the label space 
%provides for the prediction of labels. This comparison model is intended to
%quantify the contribution of the structure alone.

For all three models, particular attention was given to the settings of the 
prior parameters for the regression coefficients. These parameters implement an
important form of regularization in HSLDA. In the setting where there are no
negative labels, a Gaussian prior over the regression parameters with a
negative mean implements a prior belief that missing labels are likely to be
negative. Thus, we evaluated model performance for all three models with a
range of values for $\mu$, the mean prior parameter for regression coefficients 
($\mu\in\left\{ -3,-2.8,-2.6,\ldots,1\right\}$).

The number of topics for all models was set to 50, the prior distributions of
$p\left(\alpha\right)$, $p\left(\alpha^{\prime}\right)$, and
$p\left(\gamma\right)$ were gamma distributed with a shape parameter of 1 and a
scale parameters of 1000.

\begin{figure}[h]
\begin{center}
%\subfloat[][\label{fig:1a}]{\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
\subfigure[][]{\label{fig:1a}\includegraphics[width=.49\textwidth]{figs/clin_pred_varying_mu}}
%\subfloat[][\label{fig:1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
\subfigure[][]{\label{fig:1b}\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
\subfigure[][]{\label{fig:clinical_roc}\includegraphics[scale=0.4]{figs/ROC_comparison_leafs}}
\caption{ROC curves for out-of-sample ICD-9 code prediction from patient free-text discharge records 
(\subref{fig:1a},\subref{fig:clinical_roc}). ROC curve for out-of-sample Amazon product category predictions from 
product free-text descriptions \subref{fig:1b}. Figures \subref{fig:1a} and \subref{fig:1b} are a function 
of the prior means of the regression parameters. Figure \subref{fig:clinical_roc} is a function of auxiliary variable threshold. In all figures, solid is 
HSLDA, dashed are independent regressors + sLDA (hierarchical 
constraints on labels ignored), and dotted is HSLDA fit by running LDA first then running 
tree-conditional regressions.}
\label{fig:main_results}
\end{center}
\end{figure}

\subsection{Evaluation and Results}

%Outline:
%1. Model evaluation goal, metrics
%2. Gold standard(s), properties
%3. Method for prediction w/ details
%4. Discussion of results
%5. Connection of results with contribution and generality etc

We evaluated our model, HSLDA, against the comparison models with a focus on
predictive performance on held-out data. Prediction performance was measured
with standard metrics -- sensitiviy (true positive rate) and 1-specificity
(false positive rate).

The gold standard for comparison was derived from the testing set in each
dataset. To make the comparison as fair as possible among models, ancestors of
observed nodes in the label hierarchy were ignored, observed nodes were
considered positive and descendents of observed nodes were considered to be
negative. In particular, since the sLDA model does not enforce the hierarchical
constraints, its performance was found to be so poor on joint prediction that
it was not included. The models can be compared on a more equal footing by
considering only on the observed labels as being positive, despite the fact
that, following the hierarchical constraints, ancestors must also be positive.
Such a gold standard will likely inflate the number of false positives because
the labels applied to any particular document are usually not as complete as
they could be.  ICD-9 codes, for instance, lack sensitivity and their use as a
gold standard could lead to correctly positive predictions being labeled as
false positives\citep{Birmetal2005}. 
However, given that the label space is often large (as in our examples) it is a
moderate assumption that erroneous false positives should not skew results
significantly.

Predictive performance in HSLDA is evaluated by $p\left(y_{l,\hat{d}}\mid w_{1:N_{\hat{d}},\hat{d}}, w_{1:N_d,1:D},  y_{l\in\mathcal{L},1:D}\right)$ 
for each test document, $\hat{d}$. For efficiency, the expectation of this
probability distribution was estimated in the following way. Expectations 
of $\mathbf{\bar{z}}_d$ and $\boldsymbol{\eta}_l$ were estimated with samples
from the posterior. Using these expectations, we performed Gibbs sampling over
the hierarchy to acquire predictive samples for the documents in the test set.
The true positive rate was calculated as the average expected labeling for
gold standard positive labels. The false positive rate was calculated 
as the average expected labeling for gold standard negative labels.

Figure~\ref{fig:1a} shows the results for the clinical data. For instance, 
a prior mean for the regression of $\mu=-1.6$ yields the following performance:
the full HSLDA model had a 
true positive rate of 0.57 and a false positive rate of 0.13, the sLDA model
had a true positive rate of 0.42 and a false positive rate of 0.07, and the
HSLDA model where LDA and the regressions were fit separately had a true
positive rate of 0.39 and a false positive rate of 0.08.

These results indicate that the full HSLDA model predicts more of the the
correct labels at a cost of an increase in the number of false positives
relative to the comparison models.

Figure~\ref{fig:1b} shows the results for the retail product data. For
instance, a prior mean for the regression of $\mu=-2.2$ yields the following
performance: the full HSLDA model had a true positive rate of 0.85 
and a false positive rate of 0.30, the sLDA model had a true positive
rate of 0.78 and a false positive rate of 0.14, and the HSLDA model where
LDA and the regressions were fit separately had a true positive rate of 0.77
and a false positive rate of 0.16. These results follow a similar pattern to the clinical data.

As sensitivity and specificity can always be traded off, we examined
sensitivity for a range of values for two different parameters -- the prior
means for the regression coefficients and the threshold for the auxiliary
variables.  The goal in this analysis was to evaluate the performance of these
models subject to more or less stringent requirements for predicting positive
labels. These two parameters have important related functions in the model. The
prior mean in combination with the auxiliary variable threshold together encode
the strength of the prior belief that unobserved labels are likely to be
negative. Effectively, the prior mean applies negative pressure to the
predictions and the auxiliary variable threshold determines the cutoff. 
For each model type, separate models were fit for each value of the 
prior mean of the regression coefficients.  This is a proper Bayesian
sensitivity analysis. In contrast, to evaluate
predictive performance as a function of the auxiliary variable threshold,
a single model was fit for each model type and prediction was evaluated
based on predictive samples drawn subject to different auxiliary variable
thresholds. These methods are significantly different since the prior mean
is varied prior to inference and the auxiliary variable threshold is varied
following inference.
%However, as intended, they both highlight model performance 
%under more or less stringent requirements for predicting positive labels.

Figure \ref{fig:main_results} shows the predictive performance of HSLDA 
relative to the three comparison models as a function of the prior mean
on regression coefficients as a receiver operating characteristic (ROC) curve. 
For low values of the auxiliary variable threshold, the models predict labels
in a more sensitive and less specific manner, creating the points in the upper
right corner of the ROC curve. As the auxiliary variable threshold is
increased, the models predict in a less sensitive and more specific manner,
creating the points in the lower left hand corner of the ROC curve. For all
values of the prior mean in both datasets HSLDA outperforms sLDA with
independent regressors. In the case of HSLDA with separately trained
regression, HSLDA outperforms in the clinical dataset but performs equally well
across the board with the retail product dataset.

%\begin{figure}[t]
%%tbp] %  figure placement: here, top, bottom, or page
% \centering \includegraphics[scale=0.4]{figs/ROC_comparison_leafs} \caption{ROC curve for Out-of-sample ICD-9 code predictions.}
%\label{fig:clinical_roc} 
%\end{figure}