### 2012-11-16 Krantz-v3/Chapters/chapter1/discussion.tex

% !TEX root =  ../../Krantz_style.tex

\section{Related Work}
\label{sec:related_work}

HSLDA does not, of course, stand alone.  Models for structured labeling of bag of words data can be designed in a number of different ways.

As shown  in Section~\ref{sec:example_applications},  SLDA can be used to solve this kind of problem directly; however, doing so requires ignoring the hierarchical
dependencies amongst the labels.  Other models that incorporate LDA and supervision and so also could be used to solve this problem include
LabeledLDA~\cite{Ramage2009} and DiscLDA~\cite{DiscLDA}.  Various applications of these models to 
computer vision and document networks have been explored~\cite{wangbleifeifei08,RelationalLDA}.
None of these models, however, leverage dependency structure in the label space.

In other non-LDA-based related work, researchers have classified documents into hierarchies (a closely related task) using naive Bayes classifiers
and support vector machines. Most of this work has been demonstrated on relatively
small datasets, small label spaces, and has focused on single label classification without
a model of documents such as LDA~\cite{mccallum99building,Dumais2000HCW,Kollerilprints291,Chakrabarti1998SFS}.  
\section{Discussion}
\label{sec:discussion}

%Despite the proliferation of LDA related models in recent years, one could
%argue that the true killer app for LDA has yet to be discovered.  LDA makes
%sense for information retrieval (IR) because  the learned latent topic vectors
%serve as good low-dimensional document representations, and, a posteriori
%similarity between such vectors can be used to compute the similarity of
%documents.  Unfortunately, computational considerations render this approach
%to IR impractical in most settings.

The SLDA model family, of which HSLDA is a member, can be understood in two
different ways. One way is to see it as a family of topic models that improve
on the topic modeling performance of LDA via the inclusion of observed
supervision. An alternative, complementary way is to see it as a set of models
that can predict labels for bag-of-word data. A large diversity of problems
can be expressed as label prediction problems for bag-of-word data. A
surprisingly large amount data possess structured labels,
either hierarchically constrained or otherwise.  HSLDA directly addresses
this kind of data and works well in practice. That it
outperforms more straightforward approaches should be of interest to
practitioners.
%
%Variational Bayes has been the predominant estimation approach applied to sLDA
%models. Hierarchical probit regression makes for tractable Markov chain Monte
%Carlo SLDA inference, a benefit that should extend to other sLDA models should
%probit regression be used for response variable prediction there too.
%
%The results in Figures~\ref{fig:1a} and \ref{fig:1b} suggest that in most cases
%it is better to do full joint estimation of HSLDA.  An alternative
%interpretation of the same results is that, if one is more sensitive to the
%performance gains that result from exploiting the structure of the labels, then
%one can, in an engineering sense, get nearly as much gain in label prediction
%performance by first fitting LDA and then fitting a hierarchical probit
%regression.  There are applied settings in which this could be advantageous.

There are many kinds of problems that have the same characteristics as this: any data  that consists of free text that has been partially or
completely categorized by human editors for instance; more fully
% In this paper we focus on
any bag-of-words data that has been, at least in part, 
categorized. 
 Examples include but are not limited to webpages and curated
hierarchical directories of the same \cite{DMOZ}, product descriptions and
catalogs, (e.g.~\cite{AMAZON} as available from \cite{SNAP})
and patient records %hospital treatment transcripts
and diagnosis codes assigned to them for bookkeeping and insurance purposes.
The model we cover in this chapter shows one way to combine 
these two sources of information into a single model that allows one to
categorize new text documents automatically, suggest labels that might be
inaccurate, compute improved similarities between documents for information
retrieval purposes, and more. %The models and techniques that we develop in
%this paper are applicable to other data as well, namely, any unstructured
%representations of data that have been hierarchically classified (e.g.,~image
%catalogs with bag-of-feature representations).

%There are several challenges entailed in incorporating a hierarchy of labels
%into the model. Among them, given a large set of potential labels (often thousands), each
%instance has only a small number of labels associated to it. Furthermore, there are no
%naturally occurring negative labeling in the data, and the absence of a label
%cannot always be interpreted as a negative labeling.

%Our work operates within the framework of topic modeling. Our approach learns
%topic models of the underlying data and labeling strategies in a joint model,
%while leveraging the hierarchical structure of the labels. 
%In this chapter we focus on ``is-a'' hierarchically structured label spaces, but label spaces that are structured differently may be accommodated with relatively minor modifications.

Extensions to this work include unbounded topic cardinality variants and
relaxations to different kinds of label structure.  Unbounded topic cardinality
variants pose interesting inference challenges.  Imposing different kinds of
label structure constraints is possible within this framework, but requires relaxing some
of the assumptions we made in deriving the sampling distributions for HSLDA inference.

%We have described a mixed membership model with hierarchical supervision. We
%have demonstrated this model in the context of document modeling with
%hierarchical multi-label supervision. Such a model is appropriate in domains
%where there are hierarchical constraints among the labels such as is the case
%in an IS-A hierarchy.

%\begin{itemize}
%\item what about the nonparametric version of this?
%\item discuss the broader goal, from the beginning of search engine time, to combine categorization and free text.  this, to our knowledge, is the first principled approach to doing so
%\end{itemize}